---
title: "Bioeconomic_model_draft"
author: "Chloe Swick"
date: "2023-12-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(nloptr)
library(knitr)
library(dplyr)
library(ggplot2)
```

$L$ = legal fishing (MT)

*int stock* = initial biomass (MT)

$discount (\delta)$ = .05

$p$ = price of landed fish (per MT)

$c$ = cost of fishing (per MT)

$\alpha$ = cost of enforcement

$e_{t}$ = enforcement effort

$\gamma$= enforcement effectiveness

$K$ = carrying capacity

$r$ = intrinsic population growth rate



**Illegal Fishing**
$$
I_{t} = \frac{(p-\gamma*e)X_t}{2*c} 
$$

**Public Benefits**
$$
\pi_t = p*L_t-c * \frac{L_t^2}{X_t}-\alpha*e
$$

**Stock Equation**
$$
X_{t} = X_{t - 1} + r*X_{t - 1} - \frac{r *(X_{t - 1})^2}{K} -(L_{t-1} +I_{t-1})
$$


## Creating the Model 

```{r}
## Objective funtion

objective <- function(L, int_stock, discount, period, p, c, g, alpha, et, K, r){
  
  ##vectors 
  X <- numeric(length = 20)
I <- numeric(length = 20)
benefits <- numeric(length = 20)
  
  ##initial values 
  X[1] = int_stock
  I[1] = ((p-(g*et))*int_stock)/(2*c)  
  
  benefits[1] = (p*L[1])-(c*(L[1]^2/X[1]))-(alpha*et)
  
   print(paste(" X =", X[1], ", I =", I[1], ", benefits =", benefits[1])) 
  
  #for loop
  for (i in 2:(period+1)) {
     
    X[i] = X[i-1] + (r*X[i-1]) - (r*(X[i-1])^2)/K - (L[i-1]+I[i-1])
    
    I[i] = ((p-(g*et))*X[i])/(2*c)
    
    benefits[i] = (p*L[i]) - (c*(L[i]^2/X[i])) - (alpha*et)
   
    print(paste("Step", i, ": X =", X[i], ", I =", I[i], ", benefits =", benefits[i])) 
  }
  
  t = seq(from = 0, period)
   pv = ((1/(1 + discount))^t) * benefits
  npv = sum(pv)
  
  print(paste("L:", L))  # Print the value of L
  print(paste("npv:", npv))  # Print the value of npv
  
  return(-npv)
  
}
```


```{r}
##constraint that harvest can't be more than stock in any period

constraint <- function(L, int_stock, discount, period, p, c, g, alpha, et, K, r){
  
  #Vectors 
    X <- numeric(length = 20)
    I <- numeric(length = 20)
    const <- numeric(length = 20)
  
  ## initial states 
  X[1] = int_stock 
  I[1] = ((p-(g*et))*X[1])/(2*c)
  const[1] = (L[1]+I[1]) - X[1]
  
  ## for loop for remaing values 
  for (i in 2: (period+1)) {
     X[i] = X[i-1] + (r*X[i-1]) - (r*X[i-1]^2)/K - (L[i-1]+I[i-1])
     I[i] = ((p-(g*et))*X[i])/(2*c)
     const[i] = (L[i]+I[i]) - X[i]
    
  }
  
  return(const)
  
}
```


Andrew note on picking parameters: increasing effort should result in illegal catch going down, stock going up, and benefits should be meaningfully affected (benefits should go up, because of stock affect, reducing costs)


## Running Nlloptr

```{r}
## effort = .75 

local_opts <- list("algorithm" = "NLOPT_LN_COBYLA", xtol_rel = 1e-15)
options <- list("algorithm" = "NLOPT_LN_AUGLAG", xtol_rel = 1e-15, maxeval = 160000, "local_opts" = local_opts)

period = 19

## running nloptr 
output_e_.75 = nloptr(x0=rep(10, times=20),
                 eval_f=objective,
                 eval_g_ineq = constraint,
                 lb=rep(0, times=20),
                 opts = options,
                 int_stock= 883400, discount= .05, period= 19,
                 p= 2397, c= 15000, g= 2375, alpha= 20000000, 
                 et= .75, K= 2265128, r= .2)
                 
round(output_e_.75$solution, 3) #effort 
-output_e_.75$objective #PVNB


```

```{r}
## Effort = 1

local_opts <- list("algorithm" = "NLOPT_LN_COBYLA", xtol_rel = 1e-15)
options <- list("algorithm" = "NLOPT_LN_AUGLAG", xtol_rel = 1e-15, maxeval = 160000, "local_opts" = local_opts)

period = 19

## running nloptr 
output_e_1 = nloptr(x0=rep(10, times=20),
                 eval_f=objective,
                 eval_g_ineq = constraint,
                 lb=rep(0, times=20),
                 opts = options,
                 int_stock= 883400, discount= .05, period= 19,
                 p= 2397, c= 15000, g= 2375, alpha= 20000000, 
                 et= 1, K= 2265128, r= .2)
                 
round(output_e_1$solution, 3) #effort 
-output_e_1$objective #PVNB


```

```{r}
## effort = 0

local_opts <- list("algorithm" = "NLOPT_LN_COBYLA", xtol_rel = 1e-15)
options <- list("algorithm" = "NLOPT_LN_AUGLAG", xtol_rel = 1e-15, maxeval = 160000, "local_opts" = local_opts)

period = 19

## running nloptr 
output_e_0 = nloptr(x0=rep(10, times=20),
                 eval_f=objective,
                 eval_g_ineq = constraint,
                 lb=rep(0, times=20),
                 opts = options,
                 int_stock= 883400, discount= .05, period= 19,
                 p= 2397, c= 15000, g= 2375, alpha= 20000000, 
                 et= 0, K= 2265128, r= .2)
                 
round(output_e_0$solution, 3) #effort 
-output_e_0$objective #PVNB


```


```{r}
## effort = .5 

local_opts <- list("algorithm" = "NLOPT_LN_COBYLA", xtol_rel = 1e-15)
options <- list("algorithm" = "NLOPT_LN_AUGLAG", xtol_rel = 1e-15, maxeval = 160000, "local_opts" = local_opts)

period = 19

## running nloptr 
output_e_.5 = nloptr(x0=rep(10, times=20),
                 eval_f=objective,
                 eval_g_ineq = constraint,
                 lb=rep(0, times=20),
                 opts = options,
                 int_stock= 883400, discount= .05, period= 19,
                 p= 2397, c= 15000, g= 2375, alpha= 20000000, 
                 et= .5, K= 2265128, r= .2)
                 
round(output_e_.5$solution, 3) #effort 
-output_e_.5$objective #PVNB


```

## Making Tables

```{r}
#Making a table for e = 1

stock_a <- function(L, int_stock= 883400, discount= .05, period= 19,
                 p= 2397, c= 15000, g= 2375, alpha= 20000000, 
                 et= 1, K= 2265128, r= .2) {
  
  ##vectors 
  X <- numeric(length = 20)
I <- numeric(length = 20)
benefits <- numeric(length = 20)
  
  ##initial values 
  X[1] = int_stock
  I[1] = ((p-(g*et))*int_stock)/(2*c)  
  
  benefits[1] = (p*L[1])-(c*(L[1]^2/X[1]))-(alpha*et)
  
   print(paste(" X =", X[1], ", I =", I[1], ", benefits =", benefits[1])) 
  
  #for loop
  for (i in 2:(period+1)) {
     
    X[i] = X[i-1] + (r*X[i-1]) - (r*(X[i-1])^2)/K - (L[i-1]+I[i-1])
    
    I[i] = ((p-(g*et))*X[i])/(2*c)
    
    benefits[i] = (p*L[i]) - (c*(L[i]^2/X[i])) - (alpha*et)
   
    print(paste("Step", i, ": X =", X[i], ", I =", I[i], ", benefits =", benefits[i])) 
  }
  
  t = seq(from = 0, period)
   pv = ((1/(1 + discount))^t) * benefits
  npv = sum(pv)
  
  print(paste("L:", L))  # Print the value of L
  print(paste("npv:", npv))  # Print the value of npv
  
# This time we return a data frame with the variables we want in the table.  The value after the = indicates the column label.
  
  return(data.frame("Year"=t, "Present Value" =pv, "Stock"=X, "Harvest" =I+L, "Illegal Fishing" = I, "legal Harvest" = L))

}

# The next command is telling R to use the solution found above, choice, to evaluate the function stock and to round the values in the data frame to two places past the decimal point.  The second command says to make a table with these values. 

e_1=round(stock_a(output_e_1$solution),3)      #  2 indicates rounding

kable(e_1)
```


```{r}
#Making a table for e = 0

stock_a <- function(L, int_stock= 883400, discount= .05, period= 19,
                 p= 2397, c= 15000, g= 2375, alpha= 20000000, 
                 et= .0, K= 2265128, r= .2) {
  
  ##vectors 
  X <- numeric(length = 20)
I <- numeric(length = 20)
benefits <- numeric(length = 20)
  
  ##initial values 
  X[1] = int_stock
  I[1] = ((p-(g*et))*int_stock)/(2*c)  
  
  benefits[1] = (p*L[1])-(c*(L[1]^2/X[1]))-(alpha*et)
  
   print(paste(" X =", X[1], ", I =", I[1], ", benefits =", benefits[1])) 
  
  #for loop
  for (i in 2:(period+1)) {
     
    X[i] = X[i-1] + (r*X[i-1]) - (r*(X[i-1])^2)/K - (L[i-1]+I[i-1])
    
    I[i] = ((p-(g*et))*X[i])/(2*c)
    
    benefits[i] = (p*L[i]) - (c*(L[i]^2/X[i])) - (alpha*et)
   
    print(paste("Step", i, ": X =", X[i], ", I =", I[i], ", benefits =", benefits[i])) 
  }
  
  t = seq(from = 0, period)
   pv = ((1/(1 + discount))^t) * benefits
  npv = sum(pv)
  
  print(paste("L:", L))  # Print the value of L
  print(paste("npv:", npv))  # Print the value of npv
  
# This time we return a data frame with the variables we want in the table.  The value after the = indicates the column label.
  
  return(data.frame("Year"=t, "Present Value" =pv, "Stock"=X, "Harvest" =I+L, "Illegal Fishing" = I, "legal Harvest" = L))

}

# The next command is telling R to use the solution found above, choice, to evaluate the function stock and to round the values in the data frame to two places past the decimal point.  The second command says to make a table with these values. 

e_0=round(stock_a(output_e_0$solution),3)      #  2 indicates rounding

kable(e_0)


```


```{r}
#Making a table for e = .5 

stock_a <- function(L, int_stock= 883400, discount= .05, period= 19,
                 p= 2397, c= 15000, g= 2375, alpha= 20000000, 
                 et= .5, K= 2265128, r= .2) {
  
  ##vectors 
  X <- numeric(length = 20)
I <- numeric(length = 20)
benefits <- numeric(length = 20)
  
  ##initial values 
  X[1] = int_stock
  I[1] = ((p-(g*et))*int_stock)/(2*c)  
  
  benefits[1] = (p*L[1])-(c*(L[1]^2/X[1]))-(alpha*et)
  
   print(paste(" X =", X[1], ", I =", I[1], ", benefits =", benefits[1])) 
  
  #for loop
  for (i in 2:(period+1)) {
     
    X[i] = X[i-1] + (r*X[i-1]) - (r*(X[i-1])^2)/K - (L[i-1]+I[i-1])
    
    I[i] = ((p-(g*et))*X[i])/(2*c)
    
    benefits[i] = (p*L[i]) - (c*(L[i]^2/X[i])) - (alpha*et)
   
    print(paste("Step", i, ": X =", X[i], ", I =", I[i], ", benefits =", benefits[i])) 
  }
  
  t = seq(from = 0, period)
   pv = ((1/(1 + discount))^t) * benefits
  npv = sum(pv)
  
  print(paste("L:", L))  # Print the value of L
  print(paste("npv:", npv))  # Print the value of npv
  
# This time we return a data frame with the variables we want in the table.  The value after the = indicates the column label.
  
  return(data.frame("Year"=t, "Present Value" =pv, "Stock"=X, "Harvest" =I+L, "Illegal Fishing" = I, "legal Harvest" = L))

}

# The next command is telling R to use the solution found above, choice, to evaluate the function stock and to round the values in the data frame to two places past the decimal point.  The second command says to make a table with these values. 

e_.5=round(stock_a(output_e_.5$solution),3)      #  2 indicates rounding

kable(e_.5)


```


```{r}
#Making a table for e=. 75 

stock_a <- function(L, int_stock= 883400, discount= .05, period= 19,
                 p= 2397, c= 15000, g= 2375, alpha= 20000000, 
                 et= .75, K= 2265128, r= .2) {
  
  ##vectors 
  X <- numeric(length = 20)
I <- numeric(length = 20)
benefits <- numeric(length = 20)
  
  ##initial values 
  X[1] = int_stock
  I[1] = ((p-(g*et))*int_stock)/(2*c)  
  
  benefits[1] = (p*L[1])-(c*(L[1]^2/X[1]))-(alpha*et)
  
   print(paste(" X =", X[1], ", I =", I[1], ", benefits =", benefits[1])) 
  
  #for loop
  for (i in 2:(period+1)) {
     
    X[i] = X[i-1] + (r*X[i-1]) - (r*(X[i-1])^2)/K - (L[i-1]+I[i-1])
    
    I[i] = ((p-(g*et))*X[i])/(2*c)
    
    benefits[i] = (p*L[i]) - (c*(L[i]^2/X[i])) - (alpha*et)
   
    print(paste("Step", i, ": X =", X[i], ", I =", I[i], ", benefits =", benefits[i])) 
  }
  
  t = seq(from = 0, period)
   pv = ((1/(1 + discount))^t) * benefits
  npv = sum(pv)
  
  print(paste("L:", L))  # Print the value of L
  print(paste("npv:", npv))  # Print the value of npv
  
# This time we return a data frame with the variables we want in the table.  The value after the = indicates the column label.
  
  return(data.frame("Year"=t, "Present Value" =pv, "Stock"=X, "Harvest" =I+L, "Illegal Fishing" = I, "legal Harvest" = L))

}

# The next command is telling R to use the solution found above, choice, to evaluate the function stock and to round the values in the data frame to two places past the decimal point.  The second command says to make a table with these values. 

e_.75=round(stock_a(output_e_.75$solution),3)      #  2 indicates rounding

kable(e_.75)


```


## Putting all the data together 

```{r}
data_all <- data.frame(e_.75$Year, e_0$Stock, e_0$Illegal.Fishing, e_0$legal.Harvest, e_.5$Stock, e_.5$Illegal.Fishing, e_.5$legal.Harvest, e_.75$Stock, e_.75$Illegal.Fishing, e_.75$legal.Harvest, e_1$Stock, e_1$Illegal.Fishing, e_1$legal.Harvest)

col_names <- c("year", "stock_0", "I_0", "L_0", "stock_.5", "I_.5", "L_.5", "stock_.75", "I_.75", "L_.75", "stock_1", "I_1", "L_1")

colnames(data_all) <- col_names
```



## Figures 


```{r}
ggplot(data = e_.75, aes(x= Year, y = Illegal.Fishing)) +
  geom_line(color = 'seagreen1')+
  geom_line(aes(y=legal.Harvest), color="turquoise1") +
  geom_line(aes(y=Harvest), color="orchid1") +
   labs(title = "Harvest With 75% Enforcement Level",
       x = "Year",
       y = "Biomass (MT)") +
  theme_minimal()

```


```{r}
ggplot(data = e_1, aes(x= Year, y = Stock)) +
  geom_line(color = 'blue')+
   labs(title = "Stock with 100% Enforcement Level",
       x = "Year",
       y = "Biomass (MT)") +
  theme_minimal()

```



```{r}
## plot of illegal fishing w/ different effort 
ggplot(data = data_all, aes(x= year)) +
  geom_line(aes(y = I_0, color = 'I_0') )+
  geom_line(aes(y = I_.5, color = "I_.5") )+
  geom_line(aes(y= I_.75, color = "I_.75") )+
  geom_line(aes(y= I_1, color = "I_1") ) +
  scale_color_manual(values = c(I_0 = "blue", I_.5 = "green", I_.75 = "purple", I_1 = "yellow"), labels = c(I_0 = "0%", I_.5 = "50%", I_.75 = "75%", I_1 = "100%")) +
   labs(title = "Illegal Fishing with varying levels of enforcement",
       x = "Year",
       y = "Biomass (MT)") +
  theme_minimal() 

```


```{r}
## plot of stock with different effort 
ggplot(data = data_all, aes(x= year)) +
  geom_line(aes(y = stock_0, color = 'stock_0') )+
  geom_line(aes(y = stock_.5, color = "stock_.5") )+
  geom_line(aes(y= stock_.75, color = "stock_.75") )+
  geom_line(aes(y= stock_1, color = "stock_1") ) +
  scale_color_manual(values = c(stock_0 = "blue", stock_.5 = "green", stock_.75 = "purple", stock_1 = "yellow"), labels = c(stock_0 = "0%", stock_.5 = "50%", stock_.75 = "75%", stock_1 = "100%")) +
   labs(title = "Yellowfin Tuna stock under varying levels of enforcement",
       x = "Year",
       y = "Biomass (MT)") +
  theme_minimal() 

```







