---
title: "NRE_Final_Project"
author: "Chloe Swick"
date: "2023-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(nloptr)
library(knitr)
library(dplyr)
```


# Stage 1--> Traditional Enforcement 

## Stock Function
this function should describe how much fish stock there is (birth minus death/harvest)

from the objective function we want to describe how what the optimal TAC should be (MSY)

How far above MSY are we currently? 

Shaefer Model: 

$$X_{t+1} = X_{t} + rX_{t}(1-X_{t}/K) - qX_t E_t$$

## with no enforcement 

```{r}
#with no enforcement 
objective_no_enf <- function(I, int_stock, discount, p, c, r, K, period) {
  
  #vectors 
  stock <- vector(mode = "numeric", length = 21)
  L <- vector(mode = "numeric", length = 21)
  benefits <- vector(mode = "numeric", length = 21)
  
  #initial values
  stock[1] = int_stock
  L[1] = int_stock*.1
  benefits[1] = p*(L[1]+I[1]) - c*((L[1]+I[1])/stock[1]) 
  
  
  #for loop
  for (i in 2:(period+1))  
    {
    ### Stock dynamics
    stock[i] = stock[i-1] + r * stock[i-1] * (1 - (stock[i-1]/K)) - L[i-1] -I[i-1] 
    L[i] = stock[i-1]*.1
    benefits[i] = p*(L[i] + I[i]) - c*((L[i]+I[i])/stock[i]) 
    
  }
  
  t = seq(from = 0, period) 
  pv = ((1/(1 + discount))^t) * benefits
  npv = sum(pv)
  
  return(-npv)
  
}
```


```{r}
### Constraint: harvest < X
constraint_no_enf <- function(I, int_stock, discount, p, c, r, K, period) {
  
 stock <- vector(mode = "numeric", length = 21)
  L <- vector(mode = "numeric", length = 21)
  const <- vector(mode = "numeric", length = 21)
  
  ### Initial states
  stock[1] = int_stock
   L[1] = int_stock*.1
  const[1] = (L+I) -stock[1]
  
  ### For-loop
  for (i in 2:(21)) {
    stock[i] = stock[i - 1] + r * stock[i - 1] * (1 - (stock[i - 1] / K)) - L[i-1] - I[i - 1]
    L[i] = stock[i-1]*.1
    const[i] = (L+I) - stock[i]
  } # End for-loop
  
  return(const)
}
```


```{r}
## Use this options list for nloptr

local_opts<-list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15)
options=list("algorithm"="NLOPT_LN_AUGLAG",xtol_rel=1e-15,maxeval=16000,"local_opts"=local_opts)

period = 20

output_no_enf = nloptr(x0=rep(1, times=21),
                 eval_f= function(I) objective_no_enf(I, int_stock = 883400, 
                                                      discount=.05, p=2397, c=1000, r=.2, K=2265128, period =20),
                 eval_g_ineq = function(I) constraint_no_enf(I, int_stock = 883400, 
                                                      discount=.05, p=2397, c=1000, r=.2, K=2265128, period =20),
                 lb=rep(0, times=21),
                 opts = options,
                 )
                 
round(output_no_enf$solution, 3) #effort 
-output_no_enf$objective #PVNB

```


## with enforcement 

```{r}
objective <- function(I, int_stock, discount, p, c, ce, r, K, period, effort, alpha) {
  
  #vectors 
  stock <- vector(mode = "numeric", length = 21)
  L <- vector(mode = "numeric", length = 21)
  benefits <- vector(mode = "numeric", length = 21)
  
  #initial values
  stock[1] = int_stock
  L[1] = int_stock*.2
  benefits[1] = p*(L[1]+I[1]) - c*((L[1]+I[1])/stock[1]) - ce*(effort) - alpha*effort*I[1]
  
  
  #for loop
  for (i in 2:(period+1))  
    {
    ### Stock dynamics
    stock[i] = stock[i-1] + r * stock[i-1] * (1 - (stock[i-1]/K)) - L[i-1] -I[i-1] 
    L[i] = stock[i-1]*.2
    benefits[i] = p*(L[i] + I[i]) - c*((L[i]+I[i])/stock[i]) 
    - ce*(effort)  - alpha*effort*I[i]
    
  }
  
  t = seq(from = 0, period) 
  pv = ((1/(1 + discount))^t) * benefits
  npv = sum(pv)
  
  return(-npv)
  
}
```

## Illegal fishing / enforcement 

Model how much of total fishing is illegal fishing. this will be a function of what TAC is set at, and what enforcement/deterrence levels are

more enforcement will lead to less IUU, and get us closer to TAC

enforcement also costs money, the more money you spend on enforcement, the less you have for other things, or enforcement in other areas 

## Optimization 

how much enforcement is optimal? how do we get the highest NPV? (benefit is amount of legal fish landed, cost is cost of enforcement, want to minimize catch)

what if we set period equal to infinity? 

constraints--> 
1. catch > or equal to TAC 
2. enforcement costs < or equal to enforcement budget 

```{r}
### Constraint: harvest < X
constraint <- function(I, int_stock, discount, p, c, ce, r, K, period, effort, alpha) {
  
 stock <- vector(mode = "numeric", length = 21)
  L <- vector(mode = "numeric", length = 21)
  const <- vector(mode = "numeric", length = 21)
  
  ### Initial states
  stock[1] = int_stock
   L[1] = int_stock*.2
  const[1] = (L[1]+I[1]) -stock[1]
  
  ### For-loop
  for (i in 2:(21)) {
    stock[i] = stock[i - 1] + r * stock[i - 1] * (1 - (stock[i - 1] / K)) - L[i-1] - I[i - 1]
    L[i] = stock[i-1]*.2
    const[i] = (L[i]+I[i]) - stock[i]
  } # End for-loop
  
  return(const)
}
```



```{r}
## Use this options list for nloptr

local_opts<-list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15)
options=list("algorithm"="NLOPT_LN_AUGLAG",xtol_rel=1e-15,maxeval=16000,"local_opts"=local_opts)

period = 20

output1 = nloptr(x0=rep(1, times=21),
                 eval_f=objective,
                 eval_g_ineq = constraint,
                 lb=rep(0, times=21),
                 opts = options,
                 int_stock=883400, discount=.05, p=2397, c=1000, ce =4000, r=.2, K=2265128, 
                 period =20, effort= .5, alpha = .4)
                 
round(output1$solution, 3) #effort 
-output1$objective #PVNB

```


```{r}
#Making a table

stock_a <- function(I, int_stock=883400, discount=.05, p=2397, c=1000, ce =4000, r=.2, K=2265128, 
                 period =20, effort= .5, alpha = .4) {
  
  #vectors 
  stock <- vector(mode = "numeric", length = 21)
  L <- vector(mode = "numeric", length = 21)
  benefits <- vector(mode = "numeric", length = 21)
  
  #initial values
  stock[1] = int_stock
  L[1] = int_stock*.2
  benefits[1] = p*(L[1]+I[1]) - c*((L[1]+I[1])/stock[1]) - ce*(effort) - alpha*effort*I[1]
  
  
  
  #for loop
  for (i in 2:(period+1))  
    {
    ### Stock dynamics
    stock[i] = stock[i-1] + r * stock[i-1] * (1 - (stock[i-1]/K)) - L[i-1] -I[i-1] 
    L[i] = stock[i-1]*.2
    benefits[i] = p*(L[i] + I[i]) - c*((L[i]+I[i])/stock[i]) 
    - ce*(effort[i])  - alpha*effort[i]*I[i]
    
  }
  
  t = seq(from = 0, period) 
  pv = ((1/(1 + discount))^t) * benefits
  npv = sum(pv)
  
  
# This time we return a data frame with the variables we want in the table.  The value after the = indicates the column label.
  
  return(data.frame("Period"=t, "Present Value" =pv, "Stock"=stock, "Harvest" =I+L, "Illegal Fishing" = I))

}

# The next command is telling R to use the solution found above, choice, to evaluate the function stock and to round the values in the data frame to two places past the decimal point.  The second command says to make a table with these values. 

tbl_1a=round(stock_a(output1$solution),3)      #  2 indicates rounding

kable(tbl_1a)

```


```{r}
plot(tbl_1a$Period, tbl_1a$Stock)
```


# Stage 2--> Enforcement with Skylight 

## optimization 

questions: 
1. can TAC be reached with budget constraint? 
  if not, how effective does Skylight need to be to reach TAC? 
  if yes, how much money is saved in reaching TAC w/ Skylight 
          how much faster can TAC be reached? 
          
2. How do these results change when we have different effectiveness ratings for Skylight? 














