---
title: "NRE_Final_Project"
author: "Chloe Swick"
date: "2023-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Stage 1--> Traditional Enforcement 

## Stock Function
this function should describe how much fish stock there is (birth minus death/harvest)

from the objective function we want to describe how what the optimal TAC should be (MSY)

How far above MSY are we currently? 

```{r}

```



## Illegal fishing / enforcement 

Model how much of total fishing is illegal fishing. this will be a function of what TAC is set at, and what enforcement/deterrence levels are

more enforcement will lead to less IUU, and get us closer to TAC

enforcement also costs money, the more money you spend on enforcement, the less you have for other things, or enforcement in other areas 

## Optimization 

how much enforcement is optimal? how do we get the highest NPV? (benefit is amount of legal fish landed, cost is cost of enforcement, want to minimize catch)

constraints--> 
1. catch > or equal to TAC 
2. enforcement costs < or equal to enforcement budget 


# Stage 2--> Enforcement with Skylight 

## optimization 

questions: 
1. can TAC be reached with budget constraint? 
  if not, how effective does Skylight need to be to reach TAC? 
  if yes, how much money is saved in reaching TAC w/ Skylight 
          how much faster can TAC be reached? 
          
2. How do these results change when we have different effectiveness ratings for Skylight? 
















